### 2022-11-10 NSF2022/Description/backgroundv2.tex
BG

\subsection{Problem formulation}
Recently, we have developed a new mathematical model of
self-organization~\cite{FuQuRyYo22,fu-ryh-qua-you2022,Fu2018_SIAM}. The
model takes ideas from the phase field method and leverages boundary
integral equations to speed up computations. To formulate the problem,
consider a collection of $N_b$-many rigid bodies $U_i(t) \subset \mathbb{R}^n$
with boundary $\Gamma_i(t)$.
We call these bodies ``granules''.
The granules are immersed 
in an incompressible, zero-Reynolds number fluid
$\Omega(t) = \RR^n \setminus \cup_{i=1}^{N_b} U_i(t)$
(Figure \ref{fig:flow_map}).

@@

The key idea is to define the dynamics using the following
boundary value problem system:
\begin{alignat}{4}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, 
  &&\quad \nabla \cdot \uu = 0, && \quad \xx \in \Omega(t), \label {eqn:stokes} \\
  \label{eqn:phase}
  -\epsilon^2 \Delta \phi + W'(\phi) &= 0, && &&\quad \xx \in \Omega(t),\\
\label{eqn:noslip}        
\frac{dF_i}{dt}(\XX,t) & = \uu(\xx) 
&&=\vv_i + \omega_i \times (\xx - \aa_i), 
&&\quad \xx \in \Gamma_i(t),\\
\label{eqn:material}
\phi(\xx) &= h_i(\XX), &&  &&\quad \xx \in \Gamma_i(t),
\end{alignat}
for $i=1,\ldots,N_b$
and where $\xx = F_i(\XX,t)$.
Finally,
\begin{equation}
\label{eqn:stressbalance}
\int_{\Gamma_i} \left(\sigma  + T_i\right)\nnu \,\dif s = \mathbf{0},\quad
\int_{\Gamma_i} (\xx - \aa_i)\times \left(\sigma + T_i\right) \nnu \,\dif s = \mathbf{0}
\end{equation}
where
\begin{equation}
\label{eqn:hydro_stress}
\sigma = \mu(\nabla \uu + \nabla \uu^T) - pI,\quad 
T_i = \gamma\left[\epsilon^{-1} W'(\phi)I
  + \epsilon\left(\tfrac{1}{2}|\nabla \phi|^2I - \nabla \phi \nabla
  \phi^T\right)\right]
\end{equation}
are the hydrodynamic and phase stresses, respectively; 
$I$ is the $n\times n$ identity matrix.
For vectors $\aa, \bb \in \mathbb{R}^2$,
we follow the usual convention of defining
the ``cross product'' as
$\aa \times \bb = (\aa^{\perp} \cdot \bb) \kk$
where $\kk = (0,0,1) \in \mathbb{R}^3$ and $(a_1,a_2)^{\perp} = (-a_2,a_1)$.

@@

In the absence of external flow,
the system~\eqref{eq:RBT}--\eqref{eqn:stressbalance}
has the energy law
\begin{equation}
\label{eq:energy_law}
  \frac{d}{dt}E
  = - \int_{\Omega_t}\tfrac{1}{2}\mu|\nabla \uu + \nabla
  \uu^T|^2 \,d\xx,\quad
    E = \gamma \int_{\Omega(t)}
  \frac{\epsilon}{2} |\nabla \phi|^2 + \frac{1}{\epsilon} W(\phi) \,d\xx.
\end{equation}
Here $E$ is the energy stored in the phase transitions.
Mathematically, the energy law states that changes
in granule configuration instantaneously change the phase energy $E$.
The dynamics are such that the change energy
is lost due to hydrodynamic dissipation.
Further details and generalizations are provided in
the Specific Aim 1 section \ref{sec:specific_aim1}.

Because $E$ depends both on $\phi$ and $\Omega(t)$,
the derivation of \eqref{eq:energy_law}
is somewhat subtle and has been worked out in detail in \cite{Fu2018_SIAM}.
It goes as follows.
Using the Rayleigh Transport Theorem, 
the time derivative out $E$ yields an energy
density flux involving the phase stress $T_i$
at the boundary, plus a bulk term involving 
the Euler-Lagrange derivative of $E.$
The Euler-Lagrange derivative vanishes due to~\eqref{eqn:phase}
and the boundary flux transforms into viscous dissipation due to 
\eqref{eqn:hydro_stress} and the Stokes equations.

@@

Finally, in terms of mathematical justifications, we mention
that in the single-well case
$W(\phi) = \tfrac{1}{2}\phi^2,$ \eqref{eqn:phase} takes the form
of a screened Laplace (SL) equation $-\epsilon^2 \Delta \phi + \phi =0$.
If the material label $h_i = 1$ everywhere, then $\phi$
has a boundary layer going from $1$ at the boundary to $0$ in the bulk
and one can prove that 
\begin{equation}
\label{eqn:math_motivation}
\lim_{\epsilon \to 0} E = \frac{\gamma}{2} \int_{\partial \Omega} 1 \,dS
\end{equation}
under the assumption that $\Omega$ has a class $\mathbf{C}^2$ boundary \cite{}.
In other words, the phase energy can be viewed as an
approximation of surface energy that ``extends'' into the bulk.
In the general case, $h_i^2$ replaces $1$ in \eqref{eqn:math_motivation}.
We are confident that \eqref{eq:RBT}--\eqref{eqn:stressbalance}
is a correct description of long-range interactions between granule
surfaces.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral formulation}
\label{sec:bie}

\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} \footnotesize A finite element method
  requires meshing the entire domain (typically into triangles), while a
  BIE only requires meshing the boundary of the domain.}
\end{wrapfigure}
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs
including the Stokes equation~\eqref{eqn:stokes} and the phase
equation~\eqref{eqn:phase} in the special case of $f(u) =
\tfrac{1}{2}u^2$. Using the Stokes equation as an example, its solution
can be written as a convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega_t \rightarrow \RR^2$
is an unknown density function. The BIE reformulation has several
immediate advantages including a dimension reduction since $\ssigma$ is
defined only on $\partial \Omega_t$, and the far field boundary
condition of the PDE is automatically satisfied. The density
function $\ssigma$ is chosen to satisfy the boundary condition, and this
results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega_t,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.

>>>>

\subsection{Problem formulation}
Recently, we have developed a new mathematical model of
self-organization~\cite{FuQuRyYo22,fu-ryh-qua-you2022,Fu2018_SIAM}. The
model takes ideas from the phase field method and leverages boundary
integral equations to speed up computations. To formulate the problem,
consider a collection of $N_b$-many rigid bodies $U_i(t) \subset \mathbb{R}^n$
with boundary $\Gamma_i(t)$.
We call these bodies ``granules''.
The granules are immersed 
in an incompressible, zero-Reynolds number fluid
$\Omega_t = \RR^n \setminus \cup_{i=1}^{N_b} U_i(t)$
(Figure \ref{fig:flow_map}).

@@

The key idea is to define the dynamics using the following
boundary value problem system:
\begin{alignat}{4}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, 
  &&\quad \nabla \cdot \uu = 0, && \quad \xx \in \Omega_t, \label {eqn:stokes} \\
  \label{eqn:phase}
  -\epsilon^2 \Delta \phi + W'(\phi) &= 0, && &&\quad \xx \in \Omega_t,\\
\label{eqn:noslip}        
\frac{dF_i}{dt}(\XX,t) & = \uu(\xx) 
&&=\vv_i + \omega_i \times (\xx - \aa_i), 
&&\quad \xx \in \Gamma_i(t),\\
\label{eqn:material}
\phi(\xx) &= h_i(\XX), &&  &&\quad \xx \in \Gamma_i(t),
\end{alignat}
for $i=1,\ldots,N_b$
and where $\xx = F_i(\XX,t)$.
Finally,
\begin{equation}
\label{eqn:stressbalance}
\int_{\Gamma_i} \left(\sigma  + T_i\right)\nnu \,\dif s = \mathbf{0},\quad
\int_{\Gamma_i} (\xx - \aa_i)\times \left(\sigma + T_i\right) \nnu \,\dif s = \mathbf{0}
\end{equation}
where
\begin{equation}
\label{eqn:hydro_stress}
\sigma = \mu(\nabla \uu + \nabla \uu^T) - pI,\quad 
T_i = \gamma\left[\epsilon^{-1} W'(\phi)I
  + \epsilon\left(\tfrac{1}{2}|\nabla \phi|^2I - \nabla \phi \nabla
  \phi^T\right)\right]
\end{equation}
are the hydrodynamic and phase stresses, respectively; 
$I$ is the $n\times n$ identity matrix.
For vectors $\aa, \bb \in \mathbb{R}^2$,
we follow the usual convention of defining
the ``cross product'' as
$\aa \times \bb = (\aa^{\perp} \cdot \bb) \kk$
where $\kk = (0,0,1) \in \mathbb{R}^3$ and $(a_1,a_2)^{\perp} = (-a_2,a_1)$.

@@

In the absence of external flow,
the system~\eqref{eq:RBT}--\eqref{eqn:stressbalance}
has the energy law
\begin{equation}
\label{eq:energy_law}
  \frac{d}{dt}E
  = - \int_{\Omega_t}\tfrac{1}{2}\mu|\nabla \uu + \nabla
  \uu^T|^2 \,d\xx,\quad
    E = \gamma \int_{\Omega_t}
  \frac{\epsilon}{2} |\nabla \phi|^2 + \frac{1}{\epsilon} W(\phi) \,d\xx.
\end{equation}
Here $E$ is the energy stored in the phase transitions.
Mathematically, the energy law states that changes
in granule configuration instantaneously change the phase energy $E$.
The dynamics are such that the change energy
is lost due to hydrodynamic dissipation.
Further details and generalizations are provided in
the Specific Aim 1 section \ref{sec:specific_aim1}.

Because $E$ depends both on $\phi$ and $\Omega_t$,
the derivation of \eqref{eq:energy_law}
is somewhat subtle and has been worked out in detail in \cite{Fu2018_SIAM}.
It goes as follows.
Using the Rayleigh Transport Theorem, 
the time derivative out $E$ yields an energy
density flux involving the phase stress $T_i$
at the boundary, plus a bulk term involving 
the Euler-Lagrange derivative of $E.$
The Euler-Lagrange derivative vanishes due to~\eqref{eqn:phase}
and the boundary flux transforms into viscous dissipation due to 
\eqref{eqn:hydro_stress} and the Stokes equations.

@@

Finally, in terms of mathematical justifications, we mention
that in the single-well case
$W(\phi) = \tfrac{1}{2}\phi^2,$ \eqref{eqn:phase} takes the form
of a screened Laplace (SL) equation $-\epsilon^2 \Delta \phi + \phi =0$.
For a fixed domain $\Omega$ with class $\mathbf{C}^2$ boundary \cite{},
if the material label $h_i = 1$ everywhere, then $\phi$
has a boundary layer going from $1$ at the boundary to $0$ in the bulk
and one can prove that 
\begin{equation}
\label{eqn:math_motivation}
\lim_{\epsilon \to 0} E = \frac{\gamma}{2} \int_{\partial \Omega} 1 \,dS.
\end{equation}
In other words, the phase energy can be viewed as an
approximation of surface energy that ``extends'' into the bulk.
In the general case, $h_i^2$ replaces $1$ in \eqref{eqn:math_motivation}.
We are confident that \eqref{eq:RBT}--\eqref{eqn:stressbalance}
is a correct description of long-range interactions between granule
surfaces.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral formulation}
\label{sec:bie}
\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} \footnotesize A finite element method
  requires meshing the entire domain (typically into triangles), while a
  BIE only requires meshing the boundary of the domain.}
\end{wrapfigure}
The solution of PDEs in a domain surrounding a mobile, dense suspension
of bodies is highly nontrivial and requires accurate,
dimension-reducting discretization
that do not require continual remeshing.  
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs
including the Stokes equation~\eqref{eqn:stokes} and the phase
equation~\eqref{eqn:phase} in the special case of $f(u) =
\tfrac{1}{2}u^2$. Using the Stokes equation as an example, its solution
can be written as a convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega_t \rightarrow \RR^2$
is an unknown density function. The BIE reformulation has several
immediate advantages including a dimension reduction since $\ssigma$ is
defined only on $\partial \Omega_t$, and the far field boundary
condition of the PDE is automatically satisfied. The density
function $\ssigma$ is chosen to satisfy the boundary condition, and this
results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega_t,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.
